Simulation_EWA

### 2019-11-15 comp_model_1019.Rmd
added tabular Q learning

```{r}
naive_bayes <- function(par,data,opp_strategy_vec,return_value) {
  # par = parameter, noise in computer choice
  # dat = data subset for one participant
  # opp_strategy_vec = c("level0","level1","level2")
  # return = -2logLik 
  # or "likelihood_by_trial" vector  (for later plotting, etc)

@@

  num_strat = length(opp_strategy_vec) 
  # set up a vector to store the likelihood
  lik_opp <- matrix(0.0,num_strat, nrow(data))
  lik_hum <- matrix(0.0,nrow(data))
  br_hum <- rep(NA,num_strat)
  # Initiate prior vector as uniform
  opp_prior_vec <-rep(1/num_strat, num_strat)

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred <- NA
      lik_opp[,t] <- opp_prior_vec
      lik_hum[t] <- 1/nopts 
    } else {
      # get prediction of opponent action from CSV files
      for (strategy in opp_strategy_vec) {
        k = match(strategy,opp_strategy_vec)
        pred_opp <- as.character(filter(pred_file_opp,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
        # Given opponent action predicted, what would be human action best response for each opp strat
        br_hum[[k]] <- best_response(t_game,as.character(pred_opp))
        #cat(as.character(br_hum[[k]]), "\n")
        # Multiply prior by likelihood of observation to get posterior 
        if(as.character(data[t,"a_action"]) == pred_opp) {
          lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
        } else {
          lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
        }
      }
      # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
      # NOTE!! -> ...assumes this is unique. check
      indices <- which(br_hum %in% as.character(data[t,"h_action"]))
      #cat(as.character(indices),  "\n")

@@

naive_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  naive_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  naive_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

save(naive_bayes_modelling,file="naive_bayes_modelling.RData")

@@

>>>>

```{r}
naive_bayes <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {
  # theta = parameter, probability computer will play its best response to what it thinks human is. Truth = 90%. in [0,1].
  # Epsilone: parameter controlling noise in human choice. Probability human will deviate from its best response. In [0,1].
  # dat = data subset for one participant
  # opp_strategy_vec = vector of possible opponent strtegies, model assumes humans restrict opp strategy space to vector   c("level0","level1","level2")
  # return = -2logLik 
  # or "likelihood_by_trial" vector  (for later plotting, etc)
  # opp_mod_transfer is a boolean. True if distribution of opponent strategies is kept across games. False otherwise.

@@

  num_strat = length(opp_strategy_vec)

  #lik_opp is a vector that holds and update probability distribution of opp strategy given actions.
  lik_opp <- matrix(0.0,num_strat, nrow(data))
  # Initiate prior vector as uniform
  opp_prior_vec <-rep(1/num_strat, num_strat)

  lik_hum <- matrix(0.0,nrow(data))
  # br_hum is vector that stores best responses of human to actions taken by each level-k ai opponent at the round.
  br_hum <- rep(NA,num_strat)

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred <- NA
      lik_hum[t] <- 1/nopts
      if (opp_mod_transfer) {
        opp_prior_vec <- rep(1/num_strat, num_strat)
        lik_opp[,t] <- opp_prior_vec
        }

    # Rounds after first
    } else {
      # get prediction of opponent action from CSV files
      for (strategy in opp_strategy_vec) {
        k = match(strategy,opp_strategy_vec)
        pred_opp <- as.character(filter(pred_file_opp,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
        # Given opponent action predicted, what would be human action best response for each opp strat
        br_hum[[k]] <- best_response(t_game,as.character(pred_opp))
        #cat(as.character(br_hum[[k]]), "\n")

        # Multiply prior by likelihood of observation to get posterior. Done here ot take advantage of For loop. 
        if(as.character(data[t,"a_action"]) == pred_opp) {
          lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
        } else {
          lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
        }
      }
      # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
      indices <- which(br_hum %in% as.character(data[t,"h_action"]))
      #cat(as.character(indices),  "\n")

@@

```{r}
# 
# naive_bayes_modelling <- list()
# for(id in unique(dat$human_id)) {
#   naive_bayes_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)

#   naive_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = 0.0, upper = 1.0, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE,parallelType=1))

#   naive_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
# }
# save(naive_bayes_modelling,file="naive_bayes_modelling.RData")

# Fitting a bayesiam model with No Transfer of opp strat distribution between games 
NT_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  NT_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # NT_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = 0.0, upper = 1.0, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE,parallelType=1))

  NT_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

save(NT_bayes_modelling,file="NT_bayes_modelling.RData")

@@

## Fitting a QL with state space consisting of last round actions 
```{r}

Q_learn_states <- function(par,data,return_value){
  # beta= inverse temperature parameter in softmax choice function
  # lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # return_value=c("-2loglik","likelihood_by_trial")
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]
  gamma <- 0.9

  #Define matrix of state spaces for each game 
  G1 <- expand.grid(c("R", "P", "S"),c("R", "P", "S"))
  states_RPS <- paste0(G1$Var1,G1$Var2)

  G2 <- expand.grid(c("F", "W", "G"), c("F", "W", "G"))
  states_FWG <- paste0(G2$Var1,G2$Var2)

  G3 <- expand.grid(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
  states_NUM <- paste0(G3$Var1,G3$Var2)

  Q_vals_RPS = matrix(0.0,9,3)
  dimnames(Q_vals_RPS) = list(states_RPS, c("R", "P", "S"))

  Q_vals_FWG = matrix(0.0,9,3)
  dimnames(Q_vals_FWG) = list(states_FWG, c("F", "W", "G"))

  Q_vals_NUM = matrix(0.0,25,5)
  dimnames(Q_vals_NUM) = list(states_NUM, c("1", "2", "3", "4", "5"))

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3

    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
      state_vec <- switch(as.character(t_game),rps=states_RPS,fwg = states_FWG, numbers = states_NUM)
      lik_hum[t] <- 1/nopts
      # Randomly select prev_state and actions for first round
      state_prev  <- sample(state_vec, size = 1)
      h_act_prev <- sample(colnames(Q_vals), size =1)
      a_act_prev <- sample(colnames(Q_vals), size =1)
      reward <- 0

    } else {
      # Get reward and past human action
      reward <- as.numeric(data[t-1,"score"])
      h_act_prev <- as.character(data[t-1,"h_action"])
      a_act_prev <- as.character(data[t-1,"a_action"])
    }
    # cat(state_prev, " This is the previous state", "\n")
    new_state <- paste0(h_act_prev,a_act_prev)
    # cat(new_state,"This is new state","\n")

    # Q_learning: update rule 
    Q_vals[state_prev, a_act_prev] <- Q_vals[state_prev, a_act_prev] + alpha*( reward + gamma*max(Q_vals[new_state,])- Q_vals[state_prev, a_act_prev])

    # Assume human chooses action probabilistically using softmax on Q values
    probs <- exp(Q_vals[new_state,]/beta)/sum(exp(Q_vals[new_state,]/beta))
    #cat(probs,"\n")

    # Get actual human action and compute likelihood
    h_act <- as.character(data[t,"h_action"])
    act_index <- match(h_act, colnames(Q_vals))
    lik_hum[t] <- probs[[act_index]]

    # Update state
    state_prev <- new_state 
    #}
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)

```{r}
# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# Q_learn_states(c(0.1,0.1), data, "-2loglik")

QL_states_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_states_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_states_modelling[[id]] <- optim(c(1,0.1),fn=Q_learn_states,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")
}

save(QL_states_modelling,file="QL_states_modelling.RData")

### 2019-11-15 param_model_recovery.Rmd
added tabular Q learning

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
# Build function that will simulate a self tuning EWA player given parameter Lambda, and generate actions for 50 rounds
# of each game. 
# NOTE : haven't introduced noise in either human or computer actions.

@@

```{r}
## Generate N lambdas in chosen interval, get the corresponding datasets of actions and scores, use Self_EWA to infer Lambda from these datasets. 
EWA_self_simulation <- data.frame()
lambdas = runif(100, 0, 10)
num_rounds <- 50 
count <- 1

@@

save(EWA_self_simulation,file="EWA_self_simulation.RData")

@@

```{r}
plot(EWA_self_simulation$real_lambda,EWA_self_simulation$S_EWA_infrd_lambda)
# cor.test(EWA_self_simulation$real_lambda, EWA_self_simulation$S_EWA_infrd_lambda, method=c("pearson", "kendall", "spearman"))

# create a table with best fitting models by condition
table(EWA_self_simulation[, "condition"],c("random","Naive_Bayes","QL","EWA","S_EWA")[apply(EWA_self_simulation[,c("Random_2LL","NB_2LL","QL_2LL","EWA_2LL","S_EWA_2LL")],1,which.min)])
```
```{r}

>>>>

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
## Build function that will simulate a self tuning EWA player given parameter Lambda, and generate actions for 50 rounds
# of each game. 
```{r}

# NOTE : haven't introduced noise in either human or computer actions.

@@

## Generate 100 datasets, one for each lambda and fit the 4 models to it : NB, QL, EWA and S_EWA 
```{r}
## Generate N lambdas in chosen interval, get the corresponding datasets of actions and scores, use Self_EWA to infer Lambda from these datasets. 
EWA_self_simulation <- data.frame()
lambdas = runif(100, 0, 1)
num_rounds <- 50 
count <- 1

@@

save(EWA_self_simulation,file="EWA_low_lambda_simulation.RData")

@@

## Plotting parameter recovery scatter plots and model recovery results
```{r}

plot(EWA_self_simulation$real_lambda,EWA_self_simulation$S_EWA_infrd_lambda)
# We can evaluate the correlation between real and infered lambdas and test it for significance
cor.test(EWA_self_simulation$real_lambda, EWA_self_simulation$S_EWA_infrd_lambda, method=c("pearson", "kendall", "spearman"))

# Model recovery: create a table with best fitting models by condition
model_recov_results <- table(EWA_self_simulation[, "condition"],c("random","N_Bayes","QL","EWA","S_EWA")[apply(EWA_self_simulation[,c("Random_2LL","NB_2LL","QL_2LL","EWA_2LL","S_EWA_2LL")],1,which.min)])

 write.csv(model_recov_results,file="model_recov_results.csv",row.names = TRUE)
 kable(model_recov_results)
```
```{r}
